---
title: "Homework 15"
format: html
---

In this homework I will be working on using and understanding different R functions.

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

Import the data set
```{r}
compensation <- read.csv(here("data", "compensation.csv"))
```

1) `summary()` function gives us the mediun, mean, interquartile range, min, and max for all the numerical variables.
```{r}
summary(compensation)
```

2) `select()` allows you to choose specific columns
```{r}
select(compensation, Fruit) # select fruit column
select(compensation, -Root) #select all but root column

```


3) `slice()` allows you to choose specific rows
```{r}
slice(compensation, 2) # selects the first two rows
slice(compensation, 2:10) # selects the second to the tenth row
slice(compensation, c(2,3,10)) #selects specific rows
```


4) `filter()` allows you to select subsets of rows that are conditioned by a column
```{r}
filter(compensation, Fruit > 80) #selects all fruits that are greater than 80
filter(compensation, Fruit >80 | Fruit <20) #selects fruit that is greater than 80 and smaller than 20
lo_hi_fruit <- filter(compensation, Fruit > 80 | Fruit < 20)
```

5) `arrange()` sorts rows based on an input
```{r}
arrange(compensation, Fruit) #sorts the data frame by increasing fruit production
```


6) `mutate()` creates a new variable in the data frame
```{r}
compensation <- mutate(compensation, logFruit = log(Fruit)) #adds a new column called logfruit

head(compensation)
```


Logical operators
- `==` equals
- `!=` does not equal
- `<` less than
- `>` greater than
- `>=` greater than or equal to
- `<=` less than or equal to
- `|` or
- `&` and

You can use two functions at the same time
```{r}
select(filter(compensation, Fruit >80), Root) # shows the roots of the fruits that are greater than 80
```

Piping allows you to put one function inside another (always begin with the data frame)
```{r}
compensation %>% 
  filter(Fruit > 80) %>%
  select(Root)
# does the same task as above but we use piping to achieve it
```

7) `group_by()` tells the data set to group its contents by splitting it into groups decided by another variable
```{r}
mean.fruit <- summarise(
  group_by(compensation, Grazing),
  meanFruit = mean(Fruit)
)
# shows the average fruit for each group (grazed/ungrazed)
```

8) `mean()` finds the average for numerical data
```{r}
compensation %>% 
  group_by(Grazing) %>%
summarise(meanFruit = mean(Fruit))
# does the same thing as above but uses piping
```

9) `sd()` takes the standard deviation of a column of numerical data
```{r}
compensation %>%
  group_by(Grazing) %>%
  summarise(
    meanFruit = mean(Fruit),
    sdFruit = sd(Fruit))
# shows the mean and sd of the fruits based on grazed or ungrazed
```

10) `with()` looks in a data frame, does what its told to, then stops looking in the data frame.
```{r}
with(compensation, Fruit > 80)
```


Looks good and I especially like that you included the logical operators on lines 61 - 69.




